---
title: "SDY180: Abundance of Plasmablasts Measured by Multiparameter Flow Cytometry"
author: "Ju Yeong Kim"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 12,
  dpi = 100,
  fig.align = "center"
)
```



## Summary

This report investigate the abundance of plasmablast (and other B cell subsets) over time after vaccination with Pneumovax, Fluzone, or no vaccination (saline control group).  It can be seen on the figure below that the plasmablast subset peaks at day 7 in both vaccine groups, with a stronger and more durable response with Pneumovax. As expected, there is no clear peak in the saline group. These results are similar to those reported in Figure 6B of [Obermoser et al. (2013)](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3681204/) published as part of the original study. This vignette reproduces a report available on [the ImmuneSpace portal](https://www.immunespace.org/reports/Studies/SDY180/runReport.view?reportId=module%3ASDY180%2Freports%2Fschemas%2Fstudy%2Fdemographics%2Fplasmablast_abundance.Rmd) using `ImmPortR`.



## Load ImmPortR and other libraries

```{r libraries, cache=FALSE}
library(ImmPortR)
library(ggplot2)
library(dplyr)
```

## Reterieve the flow cytometry analyzed results

```{r connection}
fcsAnalyzed <- query_dataset("SDY180", "fcsAnalyzed")
```



## Subset the population of interest

```{r data-subset}
fcsAnalyzed <- fcsAnalyzed %>% 
  filter(grepl("Plasma", populationNameReported)) %>% 
  mutate(
    cohort = gsub("Study g", "G", armName),
    populationStatisticReported = as.double(populationStatisticReported),
    studyTimeCollected = as.factor(studyTimeCollected),
    populationStatisticReported = as.double(populationStatisticReported)
  )
```



## Compute the median

```{r data-summary}
fcsAnalyzedMedian <- fcsAnalyzed %>% 
  group_by(cohort, studyTimeCollected, populationNameReported) %>% 
  summarise(medianCellReported = median(populationStatisticReported + 1, na.rm = TRUE))
```



## Abundance of plasmablasts measured by multi-parameter flow cytometry

```{r visualization, dev='png'}
ggplot(fcsAnalyzed, aes(x = studyTimeCollected, y = populationStatisticReported + 1)) +
  geom_boxplot() +
  geom_jitter() +
  scale_y_log10() +
  facet_grid(cohort ~ populationNameReported, scale = "free") +
  xlab("Time") +
  ylab(expression(paste("Number of cells/", mu, "l"))) +
  geom_line(data = fcsAnalyzedMedian, aes(
    x = studyTimeCollected, y = medianCellReported,
    group = 1
  ), color = "black", size = 1.2) +
  labs(title = "Plasma cell abundance after vaccination")
```

